---
title: "Homework 2"
author: Jiying Han
output: github_document
---


```{r setup}
library(tidyverse)
library(readxl)
```


## Problem 1

Read the Mr. Trashwheel dataset.

```{r read_dataset}
trashwheel_df = 
  read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
            sheet = "Mr. Trash Wheel",
            range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```


Read the precipitation data! for 2018 and 2017

```{r}
precip_2018 = 
  read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet = "2018 Precipitation", 
  skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)

precip_2017 = 
  read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet = "2017 Precipitation", 
  skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

Now combine annual precipitation.

```{r}
precip_df = 
  bind_rows(precip_2018,precip_2017) %>% 
  mutate(month.name[month])
```


This dataset contains information from the Mr.Trashwheel collector in Baltimore. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month precipitation data.



## Problem 2


Read and clean the data.
```{r}
nyc_df = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>% 
  select(line:entry,vending,ada) %>% 
  mutate(entry = recode(entry, "YES" = TRUE,  "NO" = FALSE) %>% 
           as.numeric()
         )
str(nyc_df)
```


Paragraph about this dataset .
– explain briefly what variables the dataset contains, describe your data cleaning steps so far, and give the dimension (rows x columns) of the resulting dataset. Are these data tidy?
This dataset contains 


How many distinct stations are there?
```{r}
distinct_station = distinct(nyc_df,station_name, line)
```


How many stations are ADA compliant?
```{r}
station_ada = 
  tibble(
    station_name = pull(nyc_df,station_name),
    line = pull(nyc_df,line),
    ada = as.numeric(pull(nyc_df,ada))
        ) 
apa_comp_df = 
  filter(station_ada, ada == 1) 
distinct(apa_comp_df, station_name, line)
```

`r distinct(apa_comp_df, station_name, line) %>% nrow()` stations are ADA compliant

What proportion of station entrances / exits without vending allow entrance?
```{r}
entra_novend = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>% 
  select(entrance_type:vending) %>% 
  mutate(vending = recode(vending, "YES" = TRUE, "NO" = FALSE)) %>% 
  filter(vending == FALSE)
entrance_exit_novend = nrow(entra_novend)
entra_novend_exit = 
  mutate(entra_novend, entry = recode(entry,"YES" = TRUE, "NO" = FALSE )) %>% 
  filter(entry == TRUE)
entrance_noexit_novend = nrow(entra_novend_exit)
proportion = entrance_noexit_novend/entrance_exit_novend
```
The proportion is `r proportion`

Reformat data so that route number and route name are distinct variables. 
```{r}
nyc_df = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>% 
  mutate(route8 = as.character(route8)) %>% 
  mutate(route9 = as.character(route9)) %>%
  mutate(route10 = as.character(route10)) %>% 
  mutate(route11 = as.character(route11)) 

reformat_nyc_df = 
  pivot_longer(
    nyc_df, route1:route11,
    names_to = "route_name",
    values_to = "route_number"
  )
```


How many distinct stations serve the A train? Of the stations that serve the A train, how many are ADA compliant?
```{r}
reformat_nyc_df1 = 
  filter(reformat_nyc_df, route_number == "A") 
distinct(reformat_nyc_df1,station_name,line)
filter(reformat_nyc_df1, ada == TRUE ) %>% 
  distinct(station_name,line)
```

## Problem 3


First, clean the data in pols-month.csv. 
```{r pols-month}

month_df = 
  tibble(
    month = 1:12,
    month_name = month.name,
  )

pols_month_df = 
  read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  separate(mon, sep="-", into = c("year", "month", "day")) %>% 
  mutate(month = as.integer(month), year = as.integer(year), day = as.integer(day))

pols_monthname_df = 
  left_join(pols_month_df,month_df, by = "month") %>% 
    select(-month) %>% 
    rename(month = month_name) %>% 
    mutate(
      dem = prez_dem + gov_dem + sen_dem + rep_dem,
      gop = prez_gop + gov_gop + sen_gop + rep_gop,
      president = dem + gop
    ) %>% 
  select(-prez_dem, -prez_gop, -day, -dem, -gop) %>% 
  relocate(year, month) 

```




Second, clean the data in snp.
```{r snp_df}
snp_df = 
  read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
  separate(date, sep = '/', into = c("month","day", "year")) %>% 
  mutate(month = as.integer(month), year = as.integer(year), day = as.integer(day))
  
snp_monthname_df = 
  left_join(snp_df, month_df, by = "month") %>% 
  select(-month, -day) %>% 
  rename(month = month_name) %>% 
  arrange(year, month) %>% 
  relocate(year, month) 
```




Third, tidy the unemployment data so that it can be merged with the previous datasets. 
```{r umemp_df}
month2_df = 
  tibble(
    month = c("jan", "feb", "mar","apr","may","jun","jul","aug","sep","oct","nov","dec"),
    month_name = month.name
  )


umemp_df = 
  read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "variable"
               )  

umemp_monthname_df = 
  left_join(umemp_df,month2_df, by = "month") %>% 
  select(-month) %>% 
  rename(month = month_name) %>% 
  relocate(year, month)
```



Join the datasets by merging snp into pols, and merging unemployment into the result.
```{r}
final_df = 
  left_join(pols_monthname_df,snp_monthname_df, by = c("year","month")) %>% 
  left_join(umemp_monthname_df, by = c("year","month"))
```


Write a short paragraph about these datasets. Explain briefly what each dataset contained, and describe the resulting dataset (e.g. give the dimension, range of years, and names of key variables).
