---
title: "Homework 2"
author: Jiying Han
output: github_document
---


```{r setup}
library(tidyverse)
library(readxl)
```


## Problem 1

Read the Mr. Trashwheel dataset.

```{r read_dataset}
trashwheel_df = 
  read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
            sheet = "Mr. Trash Wheel",
            range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```


Read the precipitation data! for 2018 and 2017

```{r}
precip_2018 = 
  read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet = "2018 Precipitation", 
  skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)

precip_2017 = 
  read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet = "2017 Precipitation", 
  skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

Now combine annual precipitation.

```{r}
precip_df = 
  bind_rows(precip_2018,precip_2017) %>% 
  mutate(month.name[month])
```


This dataset contains information from the Mr.Trashwheel collector in Baltimore. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month precipitation data.



## Problem 2



* Read and clean the data.
```{r}
nyc_df = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>% 
  select(line:entry,vending,ada) %>% 
  mutate(entry = recode(entry, "YES" = TRUE,  "NO" = FALSE) %>% 
           as.numeric()
         )
str(nyc_df)
```



* This dataset contains `r names(nyc_df)` variables. 

\ There are `r ncol(nyc_df)`  columns and `r nrow(nyc_df)` rows are contained in this dataset. From my point of view, the original data isnt' a tidy data, for several reasons:

\ 1. In the head line, from route1-route11, they all belong to a same variable and are supposed to put in a single column.

\ 2. The value in route1:route11 are different.

\ 3. Values in _Entry_ and _Vending_ are _YES_ or _NO_, insteading of _1_ or _0_. It's might be a trouble for people to process the data.



* For clean data steps:

\ 1. I clean the data's name, which make sure that all the names are composed of lower case. It's so convinent for me to do the later process. 

\ 2.  I select interest data form the whole dataset.

\ 3. I convert the content in _entry_ column to numbers. 

\ 4. Later, I will convert columns  _route8:route11_ type to make sure that all route columns keep the same type.





```{r}
distinct_station = distinct(nyc_df,station_name, line)
```

* There are `r distinct_station %>% nrow()` distinct stations.



```{r}
station_ada = 
  tibble(
    station_name = pull(nyc_df,station_name),
    line = pull(nyc_df,line),
    ada = as.numeric(pull(nyc_df,ada))
        ) 
apa_comp_df = 
  filter(station_ada, ada == 1) 
distinct(apa_comp_df, station_name, line)
```

* There are `r distinct(apa_comp_df, station_name, line) %>% nrow()` stations are ADA compliant.



```{r}
entra_novend = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>% 
  select(entrance_type:vending) %>% 
  mutate(vending = recode(vending, "YES" = TRUE, "NO" = FALSE)) %>% 
  filter(vending == FALSE)
entrance_exit_novend = nrow(entra_novend)
entra_novend_exit = 
  mutate(entra_novend, entry = recode(entry,"YES" = TRUE, "NO" = FALSE )) %>% 
  filter(entry == TRUE)
entrance_noexit_novend = nrow(entra_novend_exit)
proportion = entrance_noexit_novend/entrance_exit_novend
```
* The proportion of station entrances / exits without vending allow entrance is `r proportion` 



* Reformat data 
```{r}
nyc_df = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>% 
  mutate(route8 = as.character(route8)) %>% 
  mutate(route9 = as.character(route9)) %>%
  mutate(route10 = as.character(route10)) %>% 
  mutate(route11 = as.character(route11)) 

reformat_nyc_df = 
  pivot_longer(
    nyc_df, route1:route11,
    names_to = "route_name",
    values_to = "route_number"
  )
```





```{r}
reformat_nyc_df1 = 
  filter(reformat_nyc_df, route_number == "A") 
distinct(reformat_nyc_df1,station_name,line)
filter(reformat_nyc_df1, ada == TRUE ) %>% 
  distinct(station_name,line)
```

* There are `r distinct(reformat_nyc_df1,station_name,line) %>% nrow()` stations serve the A train, and `r filter(reformat_nyc_df1, ada == TRUE ) %>% distinct(station_name,line) %>% nrow()` are ADA compliant.



## Problem 3


* First, clean the data in pols-month.csv
```{r pols-month}

month_df = 
  tibble(
    month = 1:12,
    month_name = month.name,
  )

pols_month_df = 
  read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  separate(mon, sep="-", into = c("year", "month", "day")) %>% 
  mutate(month = as.integer(month), year = as.integer(year), day = as.integer(day))

pols_monthname_df = 
  left_join(pols_month_df,month_df, by = "month") %>% 
    select(-month) %>% 
    rename(month = month_name) %>% 
    mutate(
      dem = prez_dem + gov_dem + sen_dem + rep_dem,
      gop = prez_gop + gov_gop + sen_gop + rep_gop,
      president = dem + gop
    ) %>% 
  select(-prez_dem, -prez_gop, -day, -dem, -gop) %>% 
  relocate(year, month) 

```




* Second, clean the data in snp.csv
```{r snp_df}
snp_df = 
  read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
  separate(date, sep = '/', into = c("month","day", "year")) %>% 
  mutate(month = as.integer(month), year = as.integer(year), day = as.integer(day))
  
snp_monthname_df = 
  left_join(snp_df, month_df, by = "month") %>% 
  select(-month, -day) %>% 
  rename(month = month_name) %>% 
  arrange(year, month) %>% 
  relocate(year, month) 
```

 


* Third, tidy the unemployment data
```{r umemp_df}
month2_df = 
  tibble(
    month = c("jan", "feb", "mar","apr","may","jun","jul","aug","sep","oct","nov","dec"),
    month_name = month.name
  )


umemp_df = 
  read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "variable"
               )  

umemp_monthname_df = 
  left_join(umemp_df,month2_df, by = "month") %>% 
  select(-month) %>% 
  rename(month = month_name) %>% 
  relocate(year, month)
```

 


* Merging snp into pols, and merging unemployment into the result
```{r}
final_df = 
  left_join(pols_monthname_df,snp_monthname_df, by = c("year","month")) %>% 
  left_join(umemp_monthname_df, by = c("year","month"))
```


* Summary for the three datasets.


There are three dataset and finally there were merged into a single dataset. 
  * For _plots_month_ dataset, 

 There are _`r names(pols_monthname_df)`_ variables. I separate the _date_ into three columns- _year_, _month_ and _day_, after that I convert the month name to number. Then add a variable called _president_. Finally select the columns and delete _prez_dem_, _prez_gop_, _day_, _dem_, _gop_. Thus, the processed pols_month dataset has `r nrow(pols_monthname_df)` rows, `r ncol(pols_monthname_df)` columns, the dimension of it is ( `r dim(pols_monthname_df)` ), the range of year is ( `r range(pols_monthname_df$year)` ).

  * For _snp_ dataset, 

I separate the date into three columns- _year_, _month_ and _day_, after that I change the type of these three columns variables, and convert the month name to number. Then select the columns and delete _month name_ and _day_. Thus, the processed dataset has _`r names( snp_monthname_df)`_ variables,  and `r nrow(snp_monthname_df)` rows, `r ncol(snp_monthname_df)` columns, the dimension of it is ( `r dim(snp_monthname_df)`) , the range of year is ( `r range(snp_monthname_df$year)` ).

  * For _unemployment_ dataset, 

There are `r names(umemp_monthname_df)` variables in processed unemployment dataset. I combine the month columns from _jan_ to _dec_, after that I use the month name to them. Then relocate _year_ and _month_ . Thus, the processed pols_month dataset has `r nrow(umemp_monthname_df)` rows , `r ncol(umemp_monthname_df)` columns, the dimension of it is ( `r dim(umemp_monthname_df)` ), the range of year is ( `r range(umemp_monthname_df$year)` ) .

  * For the resulting dataset, 

I merged above three processed datasets, and the resulting dataset has _`r names(final_df)`_ variables. It has `r nrow(final_df)` rows and `r ncol(final_df)` columns. The dimension is ( `r dim(final_df)` ). Range of year is ( `r range(final_df$year)` ).
